---
title: "HW_06"
author: "Judson Webb"
date: "February 20, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(yrbss)
```

#Exercise 1:

##Use select and filter to get a subset that shows the following columns and their conditions. Also, use the head function to show the first few rows. (Note: Columns names are in parathenses() and name of condition is in brackets[]):

###1. Females [Female] (sex) who are greater than or equal to the age of 65 (detail_age) and the method ofsuicide (short_358) was by firearms [Firearm].

###2.[Never married, single] (marital_status) males (sex), who are between the ages 20 and 25 inclusively(detail_age). (Note: If you desire, you can use between() for the condition of detail_age).

###3. People who are less than or equal to 21 years old (detail_age) and committed suicide in January(month_of_death) and their educational level (education_2003_revision) is [HS Grad] or [Bachelor’s].(Note: If you desire, you can use %in% in the condition for educational level).2 [20]
```{r, echo=TRUE}
mort <- read.csv("mortality_data.csv", header = TRUE) 

sub1 <-  select(mort, sex, detail_age, short_358)
sub1 <-  filter(sub1, sex == "Female", detail_age >= 65, short_358 == "Firearm")
head(sub1)

sub2 <- select(mort , marital_status , sex , detail_age)
sub2 <- filter(sub2,marital_status == "Never married, single", sex == "Male", detail_age>=20 & detail_age <= 25)
head(sub2)

sub3 <- select(mort, detail_age, month_of_death, education_2003_revision)
sub3 <- filter(sub3,detail_age <= 21, month_of_death == "January", 
    education_2003_revision == "HS Grad" | education_2003_revision == "Bachelor's")
head(sub3)
```

#Exercise 2:

##1. Select the columns sex, detail_age, and marital_status.

##2.Use filter to get a subset that shows the following (Note that the columns names are in parathenses()):
###a. Divorced (marital status)
###b. 50 or above (detail_age)
###c. Male (sex)
##3.Using the columns from part 1, use arrange to sort by age (detail_age) and marital status (marital_status). Also, use the head function to show the first few rows.


```{r, echo=TRUE}
ex2sub <- select(mort, sex, detail_age, marital_status)
ex2sub <- filter(ex2sub, sex == "Male", detail_age >= 50, marital_status == "Divorced")
ex2sub <- arrange(ex2sub, detail_age, marital_status)
head(ex2sub)
```

#Exercise 3:

##Please note to use the head function to see the first few rows for the following below:
###1. Select the columns that start with “s”.
###2. Select the columns that end with “x”.
###3. Select the columns that contain “age”.


```{r, echo=TRUE}
ex3sub <- select(mort, starts_with("s"))
head(ex3sub)
ex3sub <-  select(mort, ends_with("x"))
head(ex3sub)
ex3sub <-  select(mort, contains("age"))
head(ex3sub)
```

#Exercise 4:

##1. Select the detail_age column.

##2.Use transmute to add two new columns (Agem5) and (Agep5), in which you subtract and add 5 years respectively. The order of the columns is Agem5, detail_age, and then Agep5. Use the head functionto see the first few rows.

##3.Use the same conditions as part 1, but now use mutate to multiply the detail_age column by 5. Again,use the head function to see the first few rows. (Note: You are keeping the existing detail_age column,but you are changing the values in this column).

```{r, echo=TRUE}
ex4sub <-  select(mort, detail_age)
ex4sub <- transmute(ex4sub, Agem5 = detail_age-5,detail_age,Agep5 = detail_age+5)
head(ex4sub)

ex4sub <-  select(mort, detail_age)
ex4sub <- mutate(ex4sub, detail_age = detail_age*5)
head(ex4sub)
```