---
title: "HW 1 by Judson Webb"
output:
  pdf_document: default
---

#### Instructions
* Produce your assignment as a RMarkdown document
  rendered to pdf (knit to pdf) when possible. If you are having problems with this, for now we will accept the following: knit to html, click "Open in browser" in the popup window, and have your browser export it (or "print it") to pdf. Just be aware that this is a temporary allowance. You should address this ASAP.
* Also submit your Rmd file (it will not be graded but we want it for
  reference purposes).
* Show all the code (use `echo=TRUE` as option in R chunks) as
well as the results.
* 10 points per exercise.
* For this HW only, we are providing a template for you to use. In the following HW you will have to produce yours.
* See Syllabus for HW policies.

#### Data for exercises:
```{r}
data <- c(23, 89,  1, 34,
          26, 85, 24, 43,
          23, 93, 29, 45,
          32, 42, 43, NA,
          21, 54, 37, 76)
```


### Positive Integral Quantities Index Vector
* Values in the index vector must lie in the set 
{1, 2, ..., length(x)}.
* Values not in this set will produce NAs.
* Corresponding elements of the vector are selected and
  concatenated, in that order, in the result.
* The index vector can be of any length and the result
is of the same length as the index vector.

```{r, echo=TRUE}
## Your solution here...
index_vector <- (1:length(data))

```

#### Exercise 1:
Extract the elements of `data` found in positions
8, 3, 7, and 5.


```{r, echo=TRUE}
extract_p_eight <- data[8]
extract_p_three <- data[3]
extract_p_seven <- data[7]
extract_p_five <- data[5]

extract_p_eight 
extract_p_three 
extract_p_seven
extract_p_five 

```

#### Exercise 2:
Random sample and simple random sample:

First search on the Internet which is the difference
between:

* simple random sample
* random sample
(keep in mind that sometimes people confuse both namings)

(Note: the difference should be related to sampling with or without replacement).

Generate both a simple random sample and a random sample
of size 10 from `data`.

Hint: look for the help of function `sample`.
Extracting directly from `data`, a simple random sample is:
```{r}
sample(data, 10)
```

However, for didactical purposes on index vectors,
I ask you to first generate
a random index vector, and use it to extract the
chosen elements from `data`.


```{r, echo=TRUE}
srs_index <- sample((1:length(data)),10,replace = FALSE)

rs_index <- sample((1:length(data)),10,replace = TRUE)

data[srs_index]
data[rs_index]

```

#### Exercise 3:
Systematic sample:
Get a systematic sample of size `n = 5` from `data`
by extracting each value that lies every `K = N/n` elements
(where `N` is the total number of elements in `data`).

Your first element needs to be randomly determined
as a number between 1 and K.

Note: Your result needs to be a vector containing
the 5 elements that are part of your systematic sample.

```{r, echo=TRUE}
n <- 5
N <- length(data)
K <- N/n

start_index <- sample((1:K),1)
##Randomly generates the start index

index_elements <- seq(from = start_index, to = length(data), by = K)

data[index_elements]
```

### Negative Integral Quantities Index Vector
* Note: remember that negative values in the index vector specify the values to be excluded.


#### Exercise 4:
Using negative indexes, obtain a sub-vector by removing
from `data` elements in positions 3, 7, and at the end
(the position at the end must be removed in a general way,
even for vectors with different sizes).


```{r, echo=TRUE}
negative_index <- -c(3,7,length(data))
data[negative_index]
```

#### Exercise 5:
`data` has an even number of elements. Devise a general
strategy to remove only the 2 elements found in the middle.


```{r, echo=TRUE}
even_removal_index <- -c(length(data)/2,(length(data)/2)+1)
data[even_removal_index]
```

#### Exercise 6:
Obtain a vector of the differences
(second - first, third - second, ...)
using negative indexes.

Note: function `diff` does exactly this. you
can use it to verify that the result of your
solution (has to be a *one liner*) is OK:
```{r}
diff(data)
```


```{r, echo=TRUE}
data[2:length(data)]-data[1:length(data)-1]
```

### Logical Index Vectors
* Index vector must be of the *same length* as the vector
  from which elements are to be selected.
* If the length of the index vector is less than the vector,
  the index vector will be recicled with perhaps unwanted
  results.
* Values corresponding to TRUE in the index vector are
  selected.
* Values corresponding to FALSE are omitted.

```{r, echo=TRUE}
```

#### Exercise 7:
Write an R expression that will return the sum value
of 820 for the elements of `data`.

Note: using function `sum` directly yields
```{r}
sum(data)
```
as `NA`s are contagious. `sum` has the parameter
`na.rm` to remove `NA`s before performing the sum.
```{r}
sum(data, na.rm = TRUE)
```

Your solution cannot use `na.rm`, but instead
use logical indexes, and be general
(not only for this specific example).


```{r, echo=TRUE}
na_vector <- !(is.na(data))
##Produces a logic vector which holds a value of TRUE
##for each index that does not correspond to NA

sum(data[na_vector])
```

#### Exercise 8:

Write an R expression that will return the positions
of values 23 and 43 in `data`.


```{r, echo=TRUE}

tt_truth_vector <- data==23
ft_truth_vector <- data==43

which(tt_truth_vector, arr.ind = TRUE)
which(ft_truth_vector, arr.ind = TRUE)
##Produces answer based on truth vectors

```

#### Exercise 9:
Verify the "Empirical rule", that states that when data
is drawn from a Normal distribution,

* about 68% of the data is in the interval
  `(mean(x) - sd(x), mean(x) + sd(x))`
* about 95% of the data is in the interval
  `(mean(x) - 2 * sd(x), mean(x) + 2 * sd(x))`
* about 99.7% of the data is in the interval
  `(mean(x) - 3 * sd(x), mean(x) + 3 * sd(x))`

Hint: use "x <- rnorm(n)" to randomly generate
      a sample of size n from a standard
      normal distribution.
        
Use n = 1000

```{r, echo=TRUE}


sample_nine <- rnorm(1000)

check_one <- (sample_nine>= mean(sample_nine) - sd(sample_nine)) & (sample_nine <= mean(sample_nine) + sd(sample_nine))

check_two <- (sample_nine>= mean(sample_nine) - 2 * sd(sample_nine)) & (sample_nine <= mean(sample_nine) + 2 * sd(sample_nine))

check_three <- (sample_nine>= mean(sample_nine) - 3 * sd(sample_nine)) & (sample_nine <= mean(sample_nine) + 3 * sd(sample_nine))

sum(check_one)/1000
sum(check_two)/1000
sum(check_three)/1000
```
