---
title: "HW_05"
author: "Judson"
date: "February 13, 2019"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Exercise 1:

##Recreate the following plot of suicide methods (short_358) by gender (sex).:

```{r, echo=TRUE}
library(ggplot2)
library(knitr)
library(tidyverse)
library(forcats)
library(gridExtra)
library(RSQLite)
library(icd)
library(jsonlite)
library(gtable)
library(grid)
library(latex2exp)
library(gridBase)
library(nnet)
theme1 <-theme_bw()+
  theme(axis.text.x =element_text(size = 8, colour = "#6b3447",angle = 45, hjust=1),
        axis.text.y =element_text(size = 8, colour = "#6b3447",angle = 0, hjust=1),
        axis.title =element_text(size = 10, colour = "#2f2f63"),
        legend.title =element_text(size = 8, colour = "#2f2f63"),
        legend.text =element_text(size = 8, colour = "#6b3447"),
        title =element_text(size = 12, colour = "#2f2f63"),
        axis.ticks =element_line(colour = "#6b3447"),
        plot.caption =element_text(size = 8, colour = "#2f2f63"),
        plot.subtitle =element_text(size = 10, colour = "#2f2f63"))
cbPalette <-c("#CC79A7", "#D55E00", "#56B4E9", "#F0E442","#009E73", "#0072B2", "#999999", "#E69F00")
mort <- read.csv("mortality_data.csv", header = TRUE) 
```

```{r, echo=TRUE}
ggplot(data = mort) + 
   aes(x = sex, fill = short_358) +
  geom_bar(position = "fill") +
  labs(title = "Method by Gender",
       x = "Sex",
       y = "Proportion of Suicides", fill = "Method") +
  theme1 +
  scale_fill_manual(values=cbPalette)
  
```

#Exercise 2

##Create the following plot of the number of suicides utilizing miscellaneous methods. Hint: subset the overall data set for miscellaneous methods (when short_358 is "Misc.").

```{r, echo=TRUE}
miscdex <- mort[120] == "Misc."
mismort <- mort[miscdex,]
mismort<- droplevels(mismort, exclude=(anyNA(levels(mismort))))

ggplot(data = mismort) + 
   aes(icd_code_10th_desc) +
  geom_bar(color = "#d1429a",fill = "#d1429a") +  
  labs(title = "Misc. Methods",
       x = "Method",
       y = "Count") +
  theme1 +
  scale_x_discrete(name ="Methods", labels=c("Intentional self-harm by blunt object"="Blunt Object","Intentional self-harm by crashing of motor vehicle"="Motor Vehicle Crash","Intentional self-harm by drowning and submersion"="Drowning","Intentional self-harm by explosive material"="Explosion","Intentional self-harm by jumping or lying in front of moving object"="Jump In Front","Intentional self-harm by other specified means"="Misc.","Intentional self-harm by sharp object"="Sharp Object","Intentional self-harm by smoke, fire and flames"="Fire","Intentional self-harm by steam, hot vapors and hot objects"="Hot Objects","Intentional self-harm by unspecified means"="Unspecified"))

```

#Exercise 3

##Use ggplot to create a density plot for Age (age_detail)(this will be the x in the density plot).
##1. Use gaussian as the kernel.
##2. Label the x-axis "Age of Deceased" and Label the y-axis "Density Estimate".
##3. Label the main title "Age Distribution of Suicide Methods".
##4. Use facet_wrap for short_358 and set scales="free".
##5.Adjust the x-axis labels to 45 degrees and have a horizontal justification of 1 (Use theme and element_text).
```{r, echo=TRUE}
ggplot(data = mort ) + 
  stat_density(aes(detail_age), kernel ="gaussian") +  
  labs(title = "Age Distribution of Suicide Methods",
       x = "Age of Deceased",
       y = "Density Estimate") +
  theme1 +
  xlim (c(0,100)) +
  facet_wrap(~short_358,scales="free")
```

##6.Describe the modality of the distributions and provide any interesting commentary on the proportion of age in regard to suicide method.

##The charts that surprise me the least are gas poisoning and drugs, which both have peaks where people are slightly younger than 50. This makes sense, as this is around the age where you have more access to harder prescription drugs as well as automobiles and a closed door garage. Firearm suicide has a slight peak in the teens, which could be due to more violent tendencies as a result of hormones, but it peaks around 50, which seems strange to me. In general though, we see most graphs having modes around the 50s so it could just be less about the method than about the large amount of people committing suicide at this age. However, this graph is really interesting in that it has the largest right tail of any of the graphs shown, and I am unsure what to attribute this to. Asphyxiation appears to have the most obvious long-term appeal having a local minumum with a value very close to the two peaks on either side, a trend that is mirrored to a lesser extent in jumping and firearms. It's worth mentioning that these are the forms of suicide most often represented in the media, although this could be a cooincidence. Misc. is hard to make any judgements about because it's so vague, but it looks a little like the average of the other graphs.

#Exercise 4

##Use ggplot to create a bar graph for the month of the death (month_of_death) with x=short_358 and thefill=short_358 being inside the geom_bar().

##1. Label the x-axis "Method" and the y-axis "Number of Suicides".
##2. Label the main title "Suicide Methods by Month".
##3. Use facet_wrap for the month_of_death and set scales to be free in the y-axis with 4 columns.
##4.Adjust the x-axis labels to 35 degrees and have a horizontal justification of 1 (Use theme andelement_text).
##5. Use scale_fill_manual with values=cbPalette and use guides(fill=FALSE).

```{r, echo=TRUE}

theme1 <-   theme1 + theme(axis.text.x =element_text(size = 8, colour = "#6b3447",angle = 35, hjust=1))
mort$month_of_death <- fct_relevel(mort$month_of_death,c("January", "February", "March", "April","May", "June", "July", "August", "September","October", "November", "December"))
ggplot(data = mort) +
  geom_bar(aes(x=short_358,fill=short_358)) +  
  labs(title = "Suicide Methods by Month",
       x = "Method",
       y = "Number of Suicides") +
  theme1 +
  facet_wrap(~month_of_death, scales = "free_y") +
  scale_fill_manual(values=cbPalette)+
  guides(fill = FALSE)
```
##6. Does there seem to be a significant difference in suicide methods based on the month of the death?

##I'm somewhat surprised to discover that there are not that many significant differences, compared to the amount I expected.
##One thing that's noticable is that the three most common methods of suicide all drop significantly in February, the opposite of what I expected considering that I had heard suicide rates are higher during the holidays.It also looks like almost every form of suicide becomes more common in the middle of the year when compared to the first and last 2 or 3 months. However, all these trends reflect decreases and increases in the overall suicide rate, rather than in the specific methods. From this we can conclude that the month has very little influence on the method of suicide, even if it does contribute to the rate.