---
title: "HW_09"
author: "Judson Webb"
date: "March 24, 2019"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=FALSE, warning = FALSE)
```

# Problem 1: String Lengths
##1. What are the minimum, maximum, mean, and median string lengths for the mortality data column called “sex”?

```{r, echo=TRUE, cache=FALSE, warning = FALSE}
library(knitr)
library(stringr)
mort <- read.csv("mortality_data.csv", header = TRUE)
min(str_length(mort$sex))
max(str_length(mort$sex))
mean(str_length(mort$sex))
median(str_length(mort$sex))
```
##2. What are the minimum, maximum, mean, and median string lengths for the mortality data columncalled “marital_status”?
```{r, echo=TRUE, cache=FALSE, warning = FALSE}
min(str_length(mort$marital_status))
max(str_length(mort$marital_status))
mean(str_length(mort$marital_status))
median(str_length(mort$marital_status))
```
# Problem 2: Substrings and String Locations
##1. Using the head function, 
###a. Substring the mortality data column “sex” to show the first character only. (Note: Use str_sub). 
###b. Locate the string “Male” in mortality data column “sex”column. (Note: Use str_locate).
```{r, echo=TRUE, cache=FALSE, warning = FALSE}
head(str_sub(mort$sex,1,1))
head(str_locate(mort$sex,"Male"))
```
##2. Using the head function,
###a. substring the mortality data column “marital_status” to show the first three characters. (Note:Use str_sub)
###b. locate the string “Married” in mortality data column “marital_status”column. (Note: Use str_locate).
```{r, echo=TRUE, cache=FALSE, warning = FALSE}
head(str_sub(mort$marital_status,1,3))
head(str_locate(mort$marital_status,"Married"))
```
# Problem 3: String Replace All
##1. Using the head function and the first 100 rows,
###a. Replace all the occurences for “Female” with “F” in the mortality data column “sex”.
###b. Replace all the occurences for “Male” with “M” in the mortality data column “sex”.
###c. Replace all the occurences for “Female” with “F” and “Male” with “M” respectively in the mortality data column “sex”.
```{r, echo=TRUE, cache=FALSE, warning = FALSE}
head(str_replace_all(mort$sex[1:100],"Female","F"))
head(str_replace_all(mort$sex[1:100],"Male","M"))
head(str_replace_all(mort$sex[1:100],c("Female" = "F","Male" = "M")))
```
##2. Using the head function and the first 100 rows,
###a. Replace all the occurences for “Bachelor’s” with “Bachelors” in the mortality data column “education_2003_revision”.
###b. Replace all the occurences for “Unknown” with “99” in the mortality data column “education_2003_revision”.
```{r, echo=TRUE, cache=FALSE, warning = FALSE}
head(str_replace_all(mort$education_2003_revision[1:100],"Bachelor's","Bachelors"))
head(str_replace_all(mort$education_2003_revision[1:100],"Unknown","99"))
```
# Problem 4: Counting
##1. Using sum and str_count, in the first 500 rows,
###a. Count how many times “Male” appears in the mortality data column “sex”.
###b. Count how many times “Female” appears in the mortality data column “sex”.
###c. Count how many times “Bachelor’s” appears in the mortality data column “education_2003_revision”.
###d. Count how many times “Master’s” appears in the mortality data column “education_2003_revision”.
```{r, echo=TRUE, cache=FALSE, warning = FALSE}
sum(str_count(mort$sex[1:500],"Male"))
sum(str_count(mort$sex[1:500],"Female"))
sum(str_count(mort$education_2003_revision[1:500],"Bachelor's"))
sum(str_count(mort$education_2003_revision[1:500],"Master's"))
```