---
title: "HW_08"
author: "Judson Webb"
date: "March 9, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=FALSE, warning = FALSE)
```

# Problem 1: SQL Basics
##1. List the sql tables for dcon.(Note: Use dbListTables. There is only one table and it is called mortality_data).
##2. List the fields in the sql table mortality_data. (Note: Use dbListFields. It should be about 3 pageslong).
##3. Create a dataframe called mydf by selecting detail_age, sex, and marital_status and ordering by detail_age. Use the head function to show the first few rows of mydf. (Note: You are using sql to select the field names).

```{r, echo=TRUE, cache=FALSE, warning = FALSE}
library(knitr)
library(RSQLite)
```

```{r, echo=TRUE, cache=FALSE, warning = FALSE}
dcon <- dbConnect(SQLite(), dbname = "mortalitydatasqlite.sqlite")
dbListTables(dcon)
dbListFields(dcon, "mortality_data") 
res <- dbSendQuery(conn = dcon, "
SELECT detail_age, sex, marital_status
FROM mortality_data
ORDER BY detail_age;
")
mydf <- dbFetch(res, -1)
dbClearResult(res)
head(mydf)
```

# Problem 2: Filtering
##1. Use the sql database and the mortality_data sql table,
###a. Select detail_age, sex, and education_2003_revision and filter where age is greater than 20 and orderby detail_age. Use head function to show first few rows. (Note: Use ‘20’ in the where clause).
```{r, echo=TRUE, cache=FALSE, warning = FALSE}
res <- dbSendQuery(conn = dcon, "
SELECT detail_age, sex, education_2003_revision
FROM mortality_data
WHERE detail_age> '20'
ORDER BY detail_age;
")
mydf <- dbFetch(res, -1)
dbClearResult(res)
head(mydf)
```
###b. Select detail_age, sex, and education_2003_revision and filter where sex is Male and order by sex. Use head function to show first few rows.


```{r, echo=TRUE, cache=FALSE, warning = FALSE}
res <- dbSendQuery(conn = dcon, "
SELECT detail_age, sex, education_2003_revision
FROM mortality_data
WHERE sex = 'Male'
ORDER BY sex;
")
mydf <- dbFetch(res, -1)
dbClearResult(res)
head(mydf)
```
###c. Select detail_age, sex, and education_2003_revision and filter where education_2003_revision is Unknown and order by education_2003_revision. Use head function to show first few rows.
```{r, echo=TRUE, cache=FALSE, warning = FALSE}
res <- dbSendQuery(conn = dcon, "
SELECT detail_age, sex, education_2003_revision
FROM mortality_data
WHERE education_2003_revision = 'Unknown'
ORDER BY education_2003_revision;
")
mydf <- dbFetch(res, -1)
dbClearResult(res)
head(mydf)
```
##2. 
###a. Select race_new and month_of_death and filter where month_of_death is between ‘A’ and ‘J’. Order by race_new and month_of_death.
```{r, echo=TRUE, cache=FALSE, warning = FALSE}
res <- dbSendQuery(conn = dcon, "
SELECT race_new, month_of_death
FROM mortality_data
WHERE month_of_death BETWEEN 'A' AND 'J'
ORDER BY race_new,month_of_death;
")
mydf <- dbFetch(res, -1)
dbClearResult(res)
head(mydf)
```
###b. Select detail_age and day_of_week_of_death and filter where day_of_week_of_death is like ‘S%’. Order by detail age and day_of_week_of_death.
```{r, echo=TRUE, cache=FALSE, warning = FALSE}
res <- dbSendQuery(conn = dcon, "
SELECT detail_age, day_of_week_of_death
FROM mortality_data
WHERE day_of_week_of_death LIKE 'S%'
ORDER BY detail_age,day_of_week_of_death;
")
mydf <- dbFetch(res, -1)
dbClearResult(res)
head(mydf)
```
#Problem 3: Aggregate Functions
##1.Use the sql database and the sql table mortality_data and count the number of observations for short_358.
```{r, echo=TRUE, cache=FALSE, warning = FALSE}
query <- paste0("
SELECT short_358,count(*)
FROM mortality_data;")
res <- dbSendQuery(dcon, query)
dbFetch(res, -1)
dbClearResult(res)
```
##2.Use the sql database and the sql table mortality_data and count the number of observations for race_new and short 358. (Note: Group by race_new and short_358).
```{r, echo=TRUE, cache=FALSE, warning = FALSE}
query <- paste0("
SELECT race_new,short_358,count(*)
FROM mortality_data
GROUP BY race_new,short_358;")
res <- dbSendQuery(dcon, query)
dbFetch(res, -1)
dbClearResult(res)
```
##3. Use the sql database and the sql table mortality_data and count the number of observations for race_new,detail_age, and sex based on the conditions where detail_age is greater than 25, educationallevel (education_2003_revision) is a HS Grad, and the sex is Male. (Note: Use ‘25’ in the where clause).
```{r, echo=TRUE, cache=FALSE, warning = FALSE}
query <- paste0("
SELECT race_new,detail_age,sex,count(*)
FROM mortality_data
WHERE detail_age > '25' AND education_2003_revision = 'HS Grad' AND sex = 'Male'
GROUP BY race_new,detail_age,sex;")
res <- dbSendQuery(dcon, query)
head(dbFetch(res, -1))
dbClearResult(res)
```
#Problem 4: Creating, Counting, Modifying, and Deleting records from a table 
##1. Create a table agegt30, where detail_age will be greater than 30.
###a. From agegt30, select race_new and sex, and count records from this table.
###b. Assign race_new as Race and count(*) as qty (Note: Group by sex and race_new).
###c. Order by sex and race.
###d. (Note: Use ‘30’ in the where clause).
```{r, echo=TRUE, cache=FALSE, warning = FALSE}
query <- "
CREATE TABLE agegt30 AS
SELECT *
FROM mortality_data
WHERE detail_age > '30';"
dbSendQuery(dcon, query)

query <- paste0("
SELECT sex,race_new as Race,count(*) as qty
FROM agegt30
GROUP BY sex,race_new
ORDER BY race_new,sex;")
res <- dbSendQuery(dcon, query)
dbFetch(res, -1)
dbClearResult(res)

query <- "
DROP TABLE agegt30"
dbSendQuery(dcon, query)
```
##1. Create a table called agelt30, where detail_age is be less than 30.
###a. Delete records from this table where sex is Male.
###b. From agelt30, select race_new and sex, and count the records from this table.
###c. Assign race_new as Race and count(*) as qty (Note: Group by sex and race_new).
###d. (Note: Use ‘30’ in the where clause).
```{r, echo=TRUE, cache=FALSE, warning = FALSE}
query <- "
CREATE TABLE agelt30 AS
SELECT *
FROM mortality_data
WHERE detail_age < '30';"
dbSendQuery(dcon, query)

query <- "
DELETE FROM agelt30
WHERE sex = 'Male';"
dbSendQuery(dcon, query)

query <- paste0("
SELECT sex,race_new as Race,count(*) as qty
FROM agelt30
GROUP BY sex,race_new;")
res <- dbSendQuery(dcon, query)
dbFetch(res, -1)
dbClearResult(res)

query <- "
DROP TABLE agelt30"
dbSendQuery(dcon, query)
```


```{r, echo=TRUE, cache=FALSE, warning = FALSE}
dbDisconnect(dcon)
```

